Global adjoint tomography -- Model GLAD-M25

### 2019-01-16 main.tex
update

{\color{Red} Wenjie: I want to leave this part to the end, when I finished reading all the papers.}

@@

In this context, there are four key issues ASDF resolves.
\begin{itemize}
    \item Robustness and stability: the data container is developed and maintained to ensure accuracy of scientific results.
    \item Data organization: the container is self-describing. Data, including waveform, source, and station information, are organized into well-defined structures.
    \item Reproducibility: the container enables scientists to keep track of what has been done to data so that results can be reproduced.
    \item Efficiency: the format provides easy mechanisms for parallel processing.
\end{itemize}
ASDF serves as a self-contained-and-explained data container while taking full advantages of parallel computing.
In our workflow,
one ASDF file contains all the information needed for data processing, including the seismic traces, event information files, and station response files.
The related APIs are carefully designed for easy data extraction and parallel implementation.

@@

\subsubsection{Stage~III: Iterations 22--25}

At iteration~22 we added another 440 earthquakes to the database,
bringing the total to 1,480 event,
and we changed the 40--100~s surface wave weights back to their setting in Stage~I.
During this stage we switched from a nonlinear conjugate gradient optimization algorithm to an L-BFGS quasi-Newton method.

@@

In this section we calculate specific changes in misfit using all 1,480 earthquakes and an identical set of weightings and windows for three models, namely, GLAD-M25, GLAD-M15, and S362ANI combined with CRUST2.0.
Table~\ref{table:misfit_reduction_M15_M25} summarizes the changes in fit in the twelve measurement categories
between the new model, GLAD-M25, and its starting model, GLAD-M15, and between GLAD-M25 and model S362ANI combined with CRUST2.0, which was the starting model for the GLAD-M15 inversion.
In all categories we observe significant misfit reductions,
with dispersive 40--100~s surface waves exhibiting the largest improvements.
The improvements in fit per component are comparable in all period bands,
indicating that the various categories are reasonably well balanced in the assessment of misfit.

@@

Checkboard tests or tests involving a chosen known structure, e.g., a plume or slab,
are frequently used to evaluate resolution in a tomographic inversion.
Such tests are extremely expensive in FWI and adjoint tomography, because they require roughly the same computational resources as the actual inversion.
In lieu of such tests,
we performed two point-spread function analyses~\citep{fichtner2011resolution,zhu2015seismic,bozdaug2016global}.

@@

An important aspect of our inversion is that it constrains shear and compressional waves simultaneously.
In Fig.~\ref{fig:global-vs} we compare global maps of our compressional wavespeed model with P models LLNL-G3Dv3~\citep{simmons2012llnl} and GAP-P4~\citep{fukao2013subducted}.
At a depth of 100~km,
GLAD-M25 shows prominent mid-oceanic ridges, cooling of the Pacific lithosphere, and the roots of numerous continental cratons.
In the upper mantle, LLNL-G3Dv3 exhibits the largest lateral variations,
e.g., very slow compressional wavespeeds at 500~km depth below Hawaii, Samoa, and Tahiti.
In the midmantle we can see remnants of subduction beneath South America, Indonesia,
and Tonga-Kermadec.
In the deep mantle, LLNL-G3Dv3 and GAP-P4 begin to fade out, but the long-wavelength patterns,
in particular the lowerparts of the Africa and Pacific superplumes, are very similar.

@@

At depths of 100---250~km we observe high wavespeeds beneath the Central Graben in the North Sea, which is clearly separated from the EEC.
The tectonic history of the Mediterranean is complex,
involving several regional plates and multiple past and present subduction zones.
We see slab ponding in the transition zone beneath the western Mediterranean associated with slab roll-back of the Appennines-Maghrebides subduction zone~\citep{Giordano2017Magmatic}.
At shallower depths, we see low wavespeeds along its western border,
which are interpreted as back-arc extension associated with the roll-back. Compared to S362ANI$+$M and S40RTS,
our model shows much sharper features for these tectonic structures.

@@

In this section, we will visit the plumes systems on the earth.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Comment out for speed-up compile of latex file
% If not, the compilation will be slowed down quite a lot due
% to the insertion of PNG files.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
    \centering
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Afar.png}\label{fig:a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Bermuda_Canary.png}\label{fig:b}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/CapeVerde_Hoggar.png}\label{fig:c}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Iceland_Eifel.png}\label{fig:d}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Easter_Galapagos.png}\label{fig:a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Marion3_Kerguelen.png}\label{fig:g}}\\
    \caption{Vertical cross sections of perturbations of shear wave velocity Vs of plumes.}
\end{figure}

@@

In this sections, we will visit some of the well-known subduction zones.

\begin{figure}[h]
    \centering
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Aeg.png}\label{fig:a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Br.png}\label{fig:b}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ha.png}\label{fig:c}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ne.png}\label{fig:d}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Wc.png}\label{fig:a}}\\[-1pt]
    \caption{Vertical cross sections of perturbations of shear wave velocity Vs of plumes.}
\end{figure}

@@

\begin{acknowledgments}

%We want to give thanks to the Oak Ridge National Lab, for providing
%  excellent support and computational resource that facilitate our research.
%  We want to thanks IRIS and ORFEUS for providing seismic datas used in this
%  study. We want to acknowledge Hejun Zhu, Min Chen, Carl Tape and Qinya Liu,
%  for providing usefull tools and great suggetions these years.
%  We want to thanks Ebru Bozdag, for providing GLAD-M15, data and software
%  she used in the first generation model.

>>>>

@@

In this context, there are four key issues ASDF resolves.
\begin{itemize}
    \item Robustness and stability: the data container is developed and maintained to ensure accuracy of scientific results.
    \item Data organization: the container is self-describing. Data, including waveform, source, and station information, are organized into well-defined structures.
    \item Reproducibility: the container enables scientists to keep track of what has been done to data so that results can be reproduced.
    \item Efficiency: the format provides easy mechanisms for parallel processing.
\end{itemize}

ASDF serves as a self-contained-and-explained data container while taking
full advantages of parallel computing. In our workflow,
one ASDF file contains all the information needed for data processing,
including the seismic traces, event information files, and station response files.
The related APIs are carefully designed for easy data extraction and parallel implementation.

@@

\subsubsection{Stage~III: Iterations 22--25}
At iteration~22 we added another 440 earthquakes to the database,
bringing the total to 1,480 event,
and we changed the 40--100~s surface wave weights back to their
setting in Stage~I. During this stage we switched from a nonlinear
conjugate gradient optimization algorithm to an L-BFGS quasi-Newton method.

@@

In this section we calculate specific changes in misfit using all 1,480
earthquakes and an identical set of weightings and windows for three models,
namely, GLAD-M25, GLAD-M15, and S362ANI combined with CRUST2.0.
Table~\ref{table:misfit_reduction_M15_M25} summarizes the changes in fit in
the twelve measurement categories between the new model, GLAD-M25, and its
starting model, GLAD-M15, and between GLAD-M25 and model S362ANI combined
with CRUST2.0, which was the starting model for the GLAD-M15 inversion.

In all categories we observe significant misfit reductions,
with dispersive 40--100~s surface waves exhibiting the largest improvements.
The improvements in fit per component are comparable in all period bands,
indicating that the various categories are reasonably well balanced in the
assessment of misfit.

@@

Checkboard tests or tests involving a chosen known structure, e.g., a plume or slab,
are frequently used to evaluate resolution in a tomographic inversion.
Such tests are extremely expensive in FWI and adjoint tomography, because they
require roughly the same computational resources as the actual inversion.
In lieu of such tests, we performed two point-spread function analyses~
\citep{fichtner2011resolution,zhu2015seismic,bozdaug2016global}.

@@

An important aspect of our inversion is that it constrains shear and
compressional waves simultaneously. In Fig.~\ref{fig:global-vs} we
compare global maps of our compressional wavespeed model with P models
LLNL-G3Dv3~\citep{simmons2012llnl} and GAP-P4~\citep{fukao2013subducted}.
At a depth of 100~km,

GLAD-M25 shows prominent mid-oceanic ridges, cooling of the Pacific lithosphere,
and the roots of numerous continental cratons. In the upper mantle, LLNL-G3Dv3
exhibits the largest lateral variations,
e.g., very slow compressional wavespeeds at 500~km depth below Hawaii, Samoa, and Tahiti.
In the midmantle we can see remnants of subduction beneath South America, Indonesia,
and Tonga-Kermadec.
In the deep mantle, LLNL-G3Dv3 and GAP-P4 begin to fade out, but the long-wavelength patterns,
in particular the lowerparts of the Africa and Pacific superplumes, are very similar.

@@

At depths of 100---250~km we observe high wavespeeds beneath the Central Graben
in the North Sea, which is clearly separated from the EEC.
The tectonic history of the Mediterranean is complex,
involving several regional plates and multiple past and present subduction zones.
We see slab ponding in the transition zone beneath the western Mediterranean
associated with slab roll-back of the Appennines-Maghrebides subduction zone
~\citep{Giordano2017Magmatic}.
At shallower depths, we see low wavespeeds along its western border,
which are interpreted as back-arc extension associated with the roll-back.
Compared to S362ANI$+$M and S40RTS,
our model shows much sharper features for these tectonic structures.

@@

In this section, we will visit the plumes systems on the earth. Fig.\ref{fig:plumes}
shows the cross section on several plumes. In addition, we also shows TX2015 model here.

Some of the plumes origins from the CMB and extend till the surface of the Earth,
even though their formation is still under heated discussion. Even though most of
the models tends to agree near the CMB, its rising shape diffs quite a lot in the
lower and upper mantle, thus dramatically changes the dynamics of the mantle
convections. Here, we present our model together with TX2015 and SEMUCB-WM1.

Our model shows that Afar plumes orignates from the CMB, narrows its diameter
in the middle mantle and broadens again above 660km discontinuity. It behaves
very similar to TX2015 while differs a lot with SEMUCB-WM1. Despite the similarity
with TX2015, our models shows stronger slow velocity anomaly around 660km
discontinuity, possibly indicating a enhanced partial melting at that depth.

In fig.\ref{fig:plumes_b}, from left to right, we can see: 1) the Farallon
subduction slabs showing as fast anomaly; 2) the bermuda hotspot that stays
above 660km discontinuity; 3) Canary hotspot that originates from the CMT and
rise all the way up the the surface of the earth; 4) the Afar hotspot from another
angle of view.

In fig.\ref{fig:plumes_c}, from left to right, we can see: 1) the subduction slab
in South America showings as a big block of fast anomaly; 2) Cape Verde hotspot
that originates from the CMT and rise all the way up to the surface; 3) Hoggar hotspot
that has strong slow anomaly from 660km to the very surface. It is very interesting
to compare the image of Hoggar in our model with the other two, while the TX2015 only
shows slow anomaly around 660km and SEMUCB-WM1 only in the very surface.

In fig.\ref{fig:plumes_d}, plumes rises right below Icleland, reaches 660km and amplifies
its anomaly and rise to the surface. Here it may indicates another place where anmaly gets
much stronger above 660, possibly indicating the partial melting below 600km.

In fig.\ref{fig:plumes_e}, from left to right: 1) the Easter hotspot; 2) the Galapogos
hotspot; 3) the Central America subduction slabs showing as fast anomaly. It is interesting
seeing that the Easter and Galapagos may originates from the same underlying plumes, which
lies in the middle of both hotspots. The underlying plumes near the CMB belongs to the
a east extension arm of the Pacific super plume.

If fig.\ref{fig:plumes:f}, we are seeing Marion and Kerguelen hotspot, which might related
the Afar at its lower mantle part. Those hotspots and plumes are very hard to recover due
to the poor data coverage at the south hemisphere.

\begin{figure}[h]
    \centering
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Afar.png}\label{fig:plumes_a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Bermuda_Canary.png}\label{fig:plumes_b}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/CapeVerde_Hoggar.png}\label{fig:plumes_c}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Iceland_Eifel.png}\label{fig:plumes_d}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Easter_Galapagos.png}\label{fig:plumes_a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Marion3_Kerguelen.png}\label{fig:plumes_g}}\\
    \caption{Vertical cross sections of perturbations of shear wave velocity Vs of plumes. (a) Afar; (b) Bermuda (left) and Canary (mid); (c) Cape Verde (mid) and Hoggar (right); (d) Iceland (mid) and Eifel (right); (e) Easter (left) and Galapagos (right); (f) Marion (mid) and Kerguelen (right). }
    \label{fig:plumes}
\end{figure}

@@

In this section, we visit another kind of geological systems on the earth,
subduction zones. In \cite{fukao2013subducted} and \cite{van2018atlas},
they both presented the subductions systems in the details. In
figure.\ref{fig:subd} we plotted our model, GLAD-M25, together with
GAP-P4 and UU-P07.

In general, all the models are in good agreement with each other, despite
the fact that our model use waveform measurements and the other two
uses travel-time. However, there are still difference when comparing those
models in details. At shallow depth, our models shows more consistent
structure in the continental crust and craton, across all the figures we
listed in fig.\ref{fig:subd}. When it comes to the upper and middle mantle,
all the models shares a a lot in common. Difference becomes obvious at depth,
especially for subductions below 1000km. For example, in fig.\ref{fig:subd_c},
it shows the Frallon subductions. Our models shows a more continous subductions
that goes all the way the lower mantle, while the other two travel-time models
only shows the subduction slab vaguely. This comes from two factors. Our model
incorporates more data, including the whole US-Array data for this region
specifically. Second, the finite-frequency kernels have more sensitivity in
depth, when compared to the classical ray-theory based inversions.

Also, as mentioned in the previous sections, we showed our model in Vp
perturbations, since in our inverison, Vp is treated as independant
parameters, while in previous waveform models, they usually fixed the
Vp and Vs ratios.

\begin{figure}[h]
    \centering
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Aeg.png}\label{fig:subd_a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Br.png}\label{fig:subd_b}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ha.png}\label{fig:subd_c}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ne.png}\label{fig:subd_d}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Wc.png}\label{fig:subd_e}}\\[-1pt]
    \caption{Vertical cross sections of perturbations of Vp in subduction regions. (a) the Aegean anomaly; (b) South American Subduction; (c) the Hatteras anomaly; (d) the Nepal Anomaly; (e) the Wichita anomaly.}
    \label{fig:subd}
\end{figure}

@@

\newpage
\begin{acknowledgments}
%We want to give thanks to the Oak Ridge National Lab, for providing
%  excellent support and computational resource that facilitate our research.
%  We want to thanks IRIS and ORFEUS for providing seismic datas used in this
%  study. We want to acknowledge Hejun Zhu, Min Chen, Carl Tape and Qinya Liu,
%  for providing usefull tools and great suggetions these years.
%  We want to thanks Ebru Bozdag, for providing GLAD-M15, data and software
%  she used in the first generation model.
